Stochastic Variance Reduced Policy Gradient

### 2017-12-30 svrpg.tex
Preliminaries: written policy gradient subsection

%%%CUSTOM COMMANDS%%%
%Math
\newcommand{\realspace}{\mathbb R}      % realspace
\newcommand{\transpose}[1]{{#1}^\texttt{T}}
\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator*{\EV}{\mathbb{E}}
\newcommand{\EVV}[2][\ppvect \in \ppspace]{\EV_{#1}\left[{#2}\right]}
\newcommand{\norm}[2][\infty]{\left\|#2\right\|_{#1}}
\newcommand{\Dij}[2]{\frac{\partial^{2}{#1}}{\partial{#2}_i\partial{#2}_j}}
\newcommand{\de}{\,\mathrm{d}}
%RL
\newcommand{\vtheta}{\boldsymbol{\theta}}
\newcommand{\Aspace}{\mathcal{A}}
\newcommand{\Sspace}{\mathcal{S}}
\newcommand{\stationary}{d_{\mu}^{\pi_{\vtheta}}(s)}
\newcommand{\policy}{\pi_{\vtheta}(a \mid s)}
\newcommand{\pol}{\pi_{\vtheta}}
\newcommand{\score}[1]{\nabla_{#1}\log\policy}
\newcommand{\Qfun}{Q^{\pi_{\vtheta}}(s,a)}
\newcommand{\Vfun}{V^{\pi_{\vtheta}}(s)}
\newcommand{\vTheta}{\boldsymbol{\Theta}}
\newcommand{\vphi}{\boldsymbol{\phi}}
\newcommand{\gradJ}[1]{\nabla_{#1}J_\mu(\vtheta)}
\newcommand{\gradApp}[1]{\hat{\nabla}_{#1}J_{\mu}(\vtheta)}
\newcommand{\eqdef}{\mathrel{\mathop:}=}
%%%%%%

@@

\section{Preliminaries}

>>>>

%%%CUSTOM COMMANDS%%%
%Math
\newcommand{\realspace}{\mathbb R}      % realspace
\newcommand{\transpose}[1]{{#1}^\texttt{T}}
\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator*{\EV}{\mathbb{E}}
\newcommand{\EVV}[2][\ppvect \in \ppspace]{\EV_{#1}\left[{#2}\right]}
\newcommand{\norm}[2][\infty]{\left\|#2\right\|_{#1}}
\newcommand{\Dij}[2]{\frac{\partial^{2}{#1}}{\partial{#2}_i\partial{#2}_j}}
\newcommand{\de}{\,\mathrm{d}}
%RL
\newcommand{\vtheta}{\boldsymbol{\theta}}
\newcommand{\Aspace}{\mathcal{A}}
\newcommand{\Sspace}{\mathcal{S}}
\newcommand{\Tspace}{\mathcal{T}}
\newcommand{\Transition}{\mathcal{P}}
\newcommand{\Reward}{\mathcal{R}}
\newcommand{\stationary}{d_{\rho}^{\pi_{\vtheta}}(s)}
\newcommand{\policy}{\pi_{\vtheta}(a \mid s)}
\newcommand{\pol}{\pi_{\vtheta}}
\newcommand{\score}[2]{\nabla_{#1}\log\pi_{#1}(#2)}
\newcommand{\Qfun}{Q^{\pi_{\vtheta}}(s,a)}
\newcommand{\Vfun}{V^{\pi_{\vtheta}}(s)}
\newcommand{\vTheta}{\boldsymbol{\Theta}}
\newcommand{\vphi}{\boldsymbol{\phi}}
\newcommand{\gradJ}[1]{\nabla_{#1}J(\vtheta)}
\newcommand{\gradApp}[1]{\hat{\nabla}_{#1}J(\vtheta)}
\newcommand{\eqdef}{\mathrel{\mathop:}=}
\newcommand{\Dataset}{\mathcal{D}}
%%%%%%

@@

\section{Preliminaries}
In this section we provide essential background on policy gradient methods for Reinforcement Learning and variance-reduced stochastic gradient methods for finite sum optimization. In the next section we will show how these methods can be combined to solve Reinforcement Learning tasks.

\subsection{Policy Gradient for Reinforcement Learning}
A Reinforcement Learning task can be modeled with a discrete-time continuous Markov Decision Process (MDP) $\langle\Sspace,\Aspace,\Transition,\Reward,\gamma,\rho\rangle$, where $\Sspace$ is a continuous state space, $\Aspace$ is a continuous action space, $\Transition$ is a Markovian transition model, where $\Transition(s'\mid s,a)$ defines the transition density from state $s$ to $s'$ under action $a$, $\Reward$ is the reward function, where $\Reward(s,a) \in [-R,R]$ is the expected reward for state-action pair $(s,a)$ and $R$ is the maximum absolute-value reward, $\gamma\in[0,1)$ is the discount factor and $\rho$ is the initial state distribution.
The agent's behavior is modeled as a policy $\pi$, where $\pi(a\mid s)$ is the density distribution over $\Aspace$ of the action to take in state $s$.
We consider episodic tasks, \ie, tasks composed of episodes of length $H$, also called time horizon. In this setting, a trajectory $\tau$ is the sequence of states and actions $(s_0,a_0,s_1,a_1,\dots,s_{H-1},a_{H-1})$ observed by following some stationary policy in a given episode, where $s_0$ is always sampled from the initial state distribution $\rho$. With abuse of notation, we denote with $\pi(\tau)$ the density distribution induced by policy $\pi$ on the set $\Tspace$ of all possible trajectories, and with $\Reward(\tau)$ the total discounted reward provided by trajectory $\tau$:
\begin{align*}
\Reward(\tau) = \sum_{t=0}^{H-1}\gamma^t\Reward(s_t,a_t).
\end{align*}
Performance $J$ allows to rank policies in terms of expected reward:
\begin{align*}
	J(\pi) = \int_{\Tspace}\pi(\tau)\Reward(\tau)\de \tau.
\end{align*}
Solving an MDP means finding a policy $\pi^*$ maximizing $J$.
\par
Policy gradient methods restrict the search for the performance-maximizing policy over a class of parametrized policies $\Pi_{\vtheta}=\{\pol : \vtheta \in \realspace^m\}$, with the only constraint that $\pol$ is differentiable in the $m$-dimensional parameter vector $\vtheta$. We denote the performance of a parametric policy with $J(\vtheta)$ for brevity. The search for a locally optimal policy is performed by means of gradient ascent, where the gradient of $J(\vtheta)$ \wrt the policy parameters is:
\begin{align*}
\gradJ{\vtheta} = \int_{\Tspace}\pol(\tau)\score{\vtheta}{\tau}\Reward(\tau)\de \tau.
\end{align*}
We consider the on-line learning problem, where trajectories must be sampled directly by interacting with the environment. In this setting, stochastic gradient descent is typically employed: after sampling a trajectory $\tau$ or a batch of $N$ trajectories $\Dataset_{N}$,
policy parameters are updated as $\vtheta\gets\vtheta + \alpha\gradApp{\vtheta}$, where $\alpha$ is a step size and $\gradApp{\vtheta}$ is an estimate of the true gradient. The REINFORCE gradient estimator provides a simple, unbiased way of estimating the gradient:
\begin{align*}
\gradApp{\vtheta} = \frac{1}{N}\sum_{n=1}^{N}\left(\sum_{h=0}^{H-1}\score{\vtheta}{a_h^n\mid s_h^n}\right)\left(\sum_{h=0}^{H-1}\gamma^h r_h^n - b\right),
\end{align*}
where subscripts denote the time step, superscripts denote the trajectory, $r_h$ is the reward actually collected at time $h$ and b can be any baseline, provided it is constant \wrt to actions.
The G(PO)MDP gradient estimator is a refinement of REINFORCE which is subject to less variance while preserving the unbiasedness:
\begin{align*}
\gradApp{\vtheta} = \frac{1}{N}\sum_{n=1}^{N}\sum_{h=0}^{H-1}\left(\sum_{k=0}^{h}\score{\vtheta}{a_h^n\mid s_h^n}\right)\left(\gamma^h r_h^n - b\right).
\end{align*}
